---
title: "AirbnbSingapore"
author: "Margot MARCHAIS--MAURICE"
date: '2022-10-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## INTRODUCTION

Blablabla - Expliquer que le projet sera entièrement codé en SQL comme si les bases étaient chargées dans un Datawharehouse. Expliquer rapidement le business modèle d'Airbnb Inspiration pour me connecter à SQL depuis Rstudio : <https://statisticsglobe.com/run-sql-query-r> <https://statisticsglobe.com/create-connect-database-r>

### Loading R packages that will allow the connection to databases

```{r}
library(readr)
library(odbc) #Contains drivers to connect to a database
library(DBI) #Contains functions for interacting with the database
```

Set up the connection

```{r}

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
