---
title: "AirbnbSingapore"
author: "Margot MARCHAIS--MAURICE"
date: '2022-10-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## INTRODUCTION

Blablabla - Expliquer que le projet sera entièrement codé en SQL comme si les bases étaient chargées dans un Datawharehouse.
Expliquer rapidement le business modèle d'Airbnb
Inspiration pour me connecter à SQL depuis Rstudio :
https://statisticsglobe.com/run-sql-query-r
https://statisticsglobe.com/create-connect-database-r


### Loading R packages that will allow the connection to databases
```{r}
library(readr)
library(odbc) #Contains drivers to connect to a database
library(DBI) #Contains functions for interacting with the database
library(RSQL) #Generate and Process 'SQL' Queries in R
library(RSQLite) #Can create an in-memory SQL database
library(tidyverse) #Provides helpful functions and sample data to use in R
```

Set up the connection
```{r}
con <- DBI::dbConnect(drv = odbc::odbc(),
                      Driver = "driver_name",
                      Server = "server_url",
                      Database = "database_name",
                      user = "user", #optional
                      password = "password") #optional
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
